\newpage
\thispagestyle{empty}

\chapter*{Acknowledgements}
\addcontentsline{toc}{part}{Acknowledgements}
\setheader{Acknowledgements}
\label{Acknowledgements}

First and foremost, I would like to thank my promotor Prof. Vadim Cheianov and my co-promotor Prof. Koenraad Schalm for giving me the independence to carve my own path, even though it might have been slightly unconventional. I'm very thankful to Vadim for his infinite patience with my questions and for addressing them sincerely even if they might have been trivial at times. To Koenraad I'm thankful for teaching me to improve my scientific rigor and for the importance he taught me to place on communicating my results well. Thank you very much. 
\par
I'm most greatly indebted to Prof. Lars Fritz, his contribution to this work goes far beyond his supervision of the two published papers that form parts of this thesis. I'm exceedingly grateful to him for welcoming me into his group, and for his wonderful mentorship over the years. I'm also thankful to Lars for introducing me to Prof. Andrew K. Mitchell, who I greatly enjoyed collaborating with. I hope this collaboration will last for quite some time to come. 
\par
I'm also very grateful to the PhD students and postdocs that I worked with in Leiden and in Utrecht - Dima Oriekhov, Mikael Fremling and Stephan Plugge. I am also particularly grateful to the several intense discussions I had with Nicolas Chagnet over physics and on numerics in general. I would also like to acknowledge master students Alina Melissakis and Nina Doerfler whose theses I helped co-supervise- I learned quite a lot from this experience.
\par
I extend thanks next to the several friends I made along the way. I will always have a special place for my covid-housemate Alicja Dutkiewicz, and my post-covid housemate Dimitrios Krommydas, with whom I felt that family was always close by. 
\par
I also thank Floris Balm, Luca Barbera, Alexander Barnaveli, Norman Bluemel,  Meike Bos, Aleksandar Bukva, Zhenya Cheipesh, Jose Collado, Tjeerd de Leede, David Dechant, Emanuele di Salvio, Alvaro Donis-Vela, Patrick Emonts, Inci Erbilen, Matthias Flor, Ireth Garcia-Aguilar, Alice Garoffolo, Peter George, Joren Harms, Yaroslav Herasymenko, Ludwig Hoffman, Stef Koenis, Ilse Kuijf, Gal Lemut, Samadarshi Maity, Savvas Malikis, Simon Marshall, Kirsten Martens, Kamil Matuszelanski, Anna Negro, Vladimir Ohanesjan, Maialen Ortego-Larrazabal, Michal Pacholski, Adrian Perez, Stefano Polla, Kshiti Sneh Rai, Midas Ratsma, Philippe Sabella-Garnier, Katerina Saiti, Kolya Semenyakin, Fabian Sies, Tereza Vakhtel, Vova Zakharov and Masha Zakharova for their warm company, and for discussions both physics and otherwise. I hope I didn't miss anyone out! 
\par
I'm also very thankful to my friends from India - Mohit Chajjed, Mohit Dhariwal, Philip Jacob, Viraj Jangam, Daksh Khanter, Pranai Kumar, Praveen PR, Saloni Priya, Suvrat Rao, Darvin Sai Teja Reddy, Baibhaw Kumar Singh, Tulja Varun for keeping me company with frequent phone calls, especially during the pandemic. 
\par
Lastly, I thank my family - Amma, Appa, Aparna and Paati-Thatha for your support, and for being understanding of my absence from home. 
\newpage
\thispagestyle{empty}